---
title: "Transform Data"
output: github_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment="")
library(tidyverse)
# install.packages("babynames")
library(babynames)
```


# dplyr

## Your Turn 1

Alter the code to select just the `n` column:

```{r}
select(babynames, n)
```

## Quiz

Which of these is NOT a way to select the `name` and `n` columns together?

```{r}
select(babynames, -c(year, sex, prop))
select(babynames, name:n)
select(babynames, starts_with("n"))
select(babynames, ends_with("n"))
```

## Your Turn 2

Use `filter`, `babynames`, and the logical operators to find:

* All of the names where prop is greater than or equal to 0.08  
* All of the children named "Sea"  

```{r}
filter(babynames, prop >= 0.08)
filter(babynames, name == "Sea")

```

## Your Turn 3

Use Boolean operators to return only the rows that contain:

* _Boys_ named Sue  
* Names that were used by exactly 5 or 6 children in 1880  
* Names that are one of Acura, Lexus, or Yugo

```{r}
filter(babynames, sex == "M", name == "Sue")
filter(babynames, n == 5 | n == 6, year == "1880")
filter(babynames, name %in% c("Acura", "Lexus", "Yugo"))
```

## Help Me

What is the smallest value of n?
What is the largest?

```{r}
arrange(babynames, desc(n)) #Largest is 99686
arrange(babynames, n) #Smallest is 5
```

## Your Turn 4

Use `%>%` to write a sequence of functions that: 

1. Filters babynames to just the girls that were born in 2017, *then...*  
2. Selects the `name` and `n` columns, *then...*  
3. Arranges the results so that the most popular names are near the top.

```{r}
girls_2017 <- babynames %>% filter(sex == "F", year == "2017") %>% select(name, n) %>% arrange(desc(n))
#Emma was the most popular female name in 2017, with 19738 children given the name!
```

## Your Turn 5

1. Trim `babynames` to just the rows that contain **your** `name` and **your** `sex`  
2. Trim the result to just the columns that will appear in your graph (not strictly necessary, but useful practice)  
3. Plot the results as a line graph with `year` on the x axis and `prop` on the y axis

```{r}
babynames %>% 
  filter(sex == "M", name == "Juan") %>% 
  select(year, prop) %>% 
  ggplot(aes(x = year, y = prop)) +
  geom_line(color = "cyan", size = 1.12) +
  labs(title = "Popularity of the name 'Juan'", x = "Year", y = "Proportion of Males born") +
  theme_dark()
```

***

# Take aways

* Extract variables with `select()`  
* Extract cases with `filter()`  
* Arrange cases, with `arrange()`  


